\vspace*{20mm}
{
\Large\bf
\begin{center}
Acknowledgements
\end{center}
}

This thesis is the culmination of many years of work and would not
have been possible without the help and support of many people.

%%% 1. Supervisor
I would like to thank my supervisors, Professors Ian Reid and David
Murray for their guidance and patience, and good spirit throughout
the long English winters.

%%% 2. Lab
If there are any ideas of any value within this thesis, credit should
undoubtedly go to the amazing people I had the privilege to work with
at the Active Vision Lab in Oxford. In particular I would like to
thank Christopher Mei and Gabe Sibley for inspiration, encouragement,
and the most interesting coffee--time discussions I have had the
privilege of participating in. In addition, I owe much to discussions
and collaborations with Michael Osborne, Matthew Blaschko, Eric
Sommerlade, and Victor Prisacariu.

%%% 3. Funding Agency
My research was funded by a Clarendon Scholarship, for which I owe
gratitude to the wonderful people at the Clarendon Foundation and the
Oxford University Press.

%%% 4. The folks back home
Throughout this work I have enjoyed the unfailing support of my
family. To my mother, Kate, my father, Roger, and my brother, Hugh: I
owe you everything. Thank you for guiding me here, guiding me through,
and always guiding me onwards.

I would also like to thank partners--in--crime Akshat Rathi, Priya and
Preethi Vijayakumar, and Michelle Fernandez, for
making my time at Oxford what it was.
